### 2015-04-06 main.tex
Inlcuded package for bookmarks. Updated .gitignore

%
\documentclass{book}

\usepackage[utf8]{inputenc}
%\usepackage[spanish]{babel}
%\usepackage[spanish]{isodate}
\usepackage[final]{listings}
\usepackage{courier}
\usepackage{multirow}
\usepackage{enumitem}
% For links
\usepackage[bookmarks]{hyperref}
\hypersetup{colorlinks=true,allcolors=blue}

\usepackage[pdftex]{graphicx}
\usepackage[toc,page]{appendix}
\usepackage[nottoc]{tocbibind}

\def\arraystretch{1.1}

%\renewcommand{\chaptername}{Capítulo}
%\renewcommand{\figurename}{Figure}

\lstset{	
	basicstyle=\footnotesize\ttfamily,
	language=bash,
	breaklines=true,
	fontadjust=true
}

\begin{document}

\title{Multisensor Architecture for an Intersection Management System}
\vfill
\vfill
\vfill
\author{Gustavo Velasco-Hernández}
\vfill
\vfill
\vfill
\date{\today}
\maketitle

\tableofcontents
\setcounter{tocdepth}{3}

\listoffigures

\listoftables

\chapter {Introduction}
\input{ch1-introduction}

\include{ch2-mdf_ims}

\input{ch3-arch_sys}

\begin{appendices}
	\chapter {Vehicular Environment Simulators}

\end{appendices}

\bibliographystyle{abbrv}
\bibliography{bibliography}
%\bibliography{ref}

\end{document}