### 2015-06-11 ch4-laser.tex
Created new chapters files (4, 5)

% ch4-laser

\chapter{Laser-Based System}

\section{Preprocessing and Background removal}
\section{Object Recognition}
\section{Feature Extraction}
\section{Tracking and Classification}