### 2018-02-18 ch1-introduction.tex
Updated chapter 1 and 3

% Introduction
\chapter [Introduction]{Introduction}
\chaptermark{Introduction}

\section{Goals}

The main goal for this proposal is to conceive a multisensor architecture for an intersection management system and deploy it in a simulated scenario. In order to achieve this goal, these objectives were set:

\begin{itemize}
\item To make a literature review on intelligent transportation systems, focusing on developments about intersection management
\item To define system architecture and specifications, and choose a platform which fits requirements for the deployment
\item To develop and implement an intersection monitoring system based on laser and video sensors using sensor fusion techniques
\item To propose and deploy a method for incident estimation and warning generation
\item To set a test and evaluation protocol for the developments made on the choosen platform 
\end{itemize}