### 2018-06-12 main.tex
Replaced \maketitle with titlepages

%
\documentclass{book}

\usepackage[utf8]{inputenc}
%\usepackage[spanish]{babel}
%\usepackage[spanish]{isodate}
\usepackage[final]{listings}
\usepackage{courier}
\usepackage{multirow}
\usepackage{enumitem}
% For links
\usepackage[bookmarks]{hyperref}
\hypersetup{colorlinks=true,allcolors=blue}

\usepackage[pdftex]{graphicx}
\usepackage[toc,page]{appendix}
\usepackage[nottoc]{tocbibind}

\usepackage{longtable}
\usepackage{pbox}
%\usepackage[sorting=none]{biblatex}
\usepackage{gensymb}
\usepackage{epigraph}

\usepackage{todonotes}

\usepackage{titling}
\usepackage{setspace}

\def\arraystretch{1.1}

%\renewcommand{\chaptername}{Capítulo}
%\renewcommand{\figurename}{Figure}

\lstset{	
	basicstyle=\footnotesize\ttfamily,
	language=bash,
	breaklines=true,
	fontadjust=true
}

\bibliographystyle{abbrv}
%\bibliographystyle{apalike}

\hyphenation{detection camera cameras Systems}

\begin{document}

\input{title1}
\input{title2}

%\title{Multisensor Architecture for an Intersection Management System\vspace{8cm}}
%\vfill
%\vfill
%\vfill
%\author{Gustavo Velasco-Hernandez\vspace{5cm}}
%\date{\today}
%\maketitle
%
%\title{Multisensor Architecture for an Intersection Management System}
%\vfill
%\vfill
%\vfill
%\author{Gustavo Velasco-Hernandez}
%\vfill
%\vfill
%\vfill
%\date{\today}
%\maketitle

\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}%
Your abstract goes here...
...

\chapter*{Acknowledgements}%
\addcontentsline{toc}{chapter}{Acknowledgements}%

I would like to thank to all people that in some way or another BLABLABLA

\tableofcontents
\setcounter{tocdepth}{3}

\listoffigures

\listoftables

\input{ch1-introduction}

\include{ch2-mdf_ims}

\include{ch3-arch_sys}

\include{ch4-results}

\include{ch5-conclusions}

\begin{appendices}
	\chapter {Vehicular Environment Simulators}

\end{appendices}

\bibliography{bibliography}
%\bibliography{ref}

\end{document}