### 2018-06-21 ch4-results.tex

\chapter [Deployment and Results]{Deployment and Results}

\section{Experimental Setup}

The selected dataset for testing the system is the one from POSSi project (section \ref{possi_ds}). It contains 6 laser-scanners raw readings and a video from a camera located over the intersection. The ground truth data used for validation is the vehicle count over 3 of the 5 legs of the intersection, taking into account the time at which a vehicle appears. This groundtruth data was generated manually.

\section{Test Configurations}
In order to test the proposed architecture, two different configurations have been selected. The first configuration is using just the camera information. The second configuration is based on multiple laser sensors along with the camera.

Each configuration consists of a set of processing blocks (as described in section \ref{proc_blocks}) connected, aiming to take data from  lasers and cameras, to produce an output of higher level. In this case, vehicle counting is the metric used for evaluation.

The following table lists all used blocks and assigns an ID for each one. Those IDs are used in the graph description in its own section. Also, bold nodes and connections indicate multiple instance of the same element.

\begin{table}[ht!]
\footnotesize
\centering
\begin{tabular}{|c | c| c|}
\hline
\textbf{Block ID} & \textbf{Name} \\
\hline
A & laser\_publisher \\
\hline
B & laser\_bg\_remover \\
\hline
C & laser\_pol2cart \\
\hline
D & laser\_cart\_merge \\
\hline
E & points2clusters \\
\hline
F & clusters\_veh\_counter \\
\hline
G & veh\_counter\_merge \\
\hline
H & clusters2occ\_level \\
\hline
I & occ\_level\_merge \\
\hline
J & laser\_pol2cart \\
\hline
K & points2occgrid \\
\hline
L & occgrid\_merge \\
\hline
M & occgrid\_leg\_counter \\
\hline
N & occgrid2occ\_level \\
\hline
O & camera\_publisher \\
\hline
P & camera\_bg\_remover \\
\hline
Q & camera\_blobs \\
\hline
R & camera\_blobs2occgrid \\
\hline
\end{tabular}
\caption{Description of processing blocks used in test configurations}
\label{desc_test_config}
\end{table}

\subsection{Case 1: Single Camera}

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.7]{fig/4/test_configuration1.pdf}
\caption{Single laser configuration}
\label{tconf1}
\end{figure}

TODO

\subsection{Case 2: Multiple Lasers}

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.7]{fig/4/test_configuration2.pdf}
\caption{Multiple lasers configuration}
\label{tconf2}
\end{figure}

TODO

%\subsection{Case 3: Multiple Lasers and camera}
%
%\begin{figure}[ht!]
%\centering
%\includegraphics[scale=0.7]{fig/4/test_configuration3.pdf}
%\caption{Multiple lasers and camera configuration}
%\label{tconf3}
%\end{figure}

TODO

\subsection{Results and Comparison}

TODO