### 2018-06-21 main.tex

%
\documentclass{book}

\usepackage[utf8]{inputenc}
%\usepackage[spanish]{babel}
%\usepackage[spanish]{isodate}
\usepackage[final]{listings}
\usepackage{courier}
\usepackage{multirow}
\usepackage{enumitem}
% For links
\usepackage[bookmarks]{hyperref}
\hypersetup{colorlinks=true,allcolors=blue}

\usepackage[pdftex]{graphicx}
\usepackage[toc,page]{appendix}
\usepackage[nottoc]{tocbibind}

\usepackage{longtable}
\usepackage{pbox}
%\usepackage[sorting=none]{biblatex}
\usepackage{gensymb}
\usepackage{epigraph}

\usepackage{todonotes}

\usepackage{titling}
\usepackage{setspace}

\def\arraystretch{1.1}

%\renewcommand{\chaptername}{Capítulo}
%\renewcommand{\figurename}{Figure}

\lstset{	
	basicstyle=\footnotesize\ttfamily,
	language=bash,
	breaklines=true,
	fontadjust=true
}

\bibliographystyle{abbrv}
%\bibliographystyle{apalike}

\hyphenation{detection camera cameras Systems}

\begin{document}

\input{title1}
\input{title2}

%\title{Multisensor Architecture for an Intersection Management System\vspace{8cm}}
%\vfill
%\vfill
%\vfill
%\author{Gustavo Velasco-Hernandez\vspace{5cm}}
%\date{\today}
%\maketitle
%
%\title{Multisensor Architecture for an Intersection Management System}
%\vfill
%\vfill
%\vfill
%\author{Gustavo Velasco-Hernandez}
%\vfill
%\vfill
%\vfill
%\date{\today}
%\maketitle

\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}%
Intersections or juntions are critical points in transportation systems due to their dynamic nature, making them prone to safety or efficiency problems. For this reason the use of technology in monitoring intersections have been increasing over the last years. After a comprehensive review on this topic, a  multisensor architecture for an intersection management system is conceived, having as main features its scalability and modularity. This architecture is composed by 4 main elements: communication structure, data model, information structure and processing blocks. An example implementation of the architecture is made using the publisher/subscriber approach along with different tools and technologies that allow the whole system to fulfill the proposed criteria, to be scalable and modular. After testing the system using a dataset containing data from camera sensors and lasers sensors, some comments about its results, future developments and integration with current needs and new trends, are discussed.

%\chapter*{Acknowledgements}%
%\addcontentsline{toc}{chapter}{Acknowledgements}%

%I would like to thank to all people that in some way or another BLABLABLA

\tableofcontents
\setcounter{tocdepth}{3}

\listoffigures

\listoftables

\input{ch1-introduction}

\include{ch2-mdf_ims}

\include{ch3-arch_sys}

\include{ch4-results}

\include{ch5-conclusions}

%\begin{appendices}
%	\chapter {Vehicular Environment Simulators}
%\end{appendices}

\bibliography{bibliography}
%\bibliography{ref}

\end{document}