### 2018-06-21 title2.tex

\begin{titlepage}
    \begin{center}
        \vspace*{1cm}
        \Huge        
        Multisensor Architecture for an Intersection Management System

        \vspace{3.5cm}
        \LARGE
        Gustavo Adolfo Velasco-Hernandez

        \vfill
        \large

        A thesis presented in partial fulfillment of the Requirement for the degree of Master of Science in Engineering with emphasis on Electronics.

        \vspace{2cm}

        Advisor: Prof. Eduardo Caicedo Bravo

		\vspace{2cm}

        %\includegraphics[width=0.4\textwidth]{university}
        Perception and Intelligent Systems Research Group \\
        School of Electrical and Electronics Engineering\\
        Universidad del Valle\\
        Colombia\\
        \today

    \end{center}
\end{titlepage}