### 2018-07-31 last_version.tex

\begin{titlepage}
    \begin{center}
%        \vspace*{1cm}
%        \Huge        
%        Multisensor Architecture for an Intersection Management System
%   
%        \vspace{3.5cm}
%        \LARGE
%        Gustavo Adolfo Velasco-Hernandez
%        
%        \vfill
%        \large
%		        
%        
%        A thesis presented in partial fulfillment of the Requirement for the degree of Master of Science in Engineering with emphasis on Electronics.
%        
%        \vspace{2cm}
%        
%        Advisor: Prof. Eduardo Caicedo Bravo
%		
%		\vspace{2cm}
%        
%        %\includegraphics[width=0.4\textwidth]{university}
%        Perception and Intelligent Systems Research Group \\
%        School of Electrical and Electronics Engineering\\
%        Universidad del Valle\\
%        Colombia\\
%        \today
%\vfill

Version: 0.8 \\
\vspace{2cm}
An updated version of this report can be downloaded from: \\
\href{https://github.com/gustavovelascoh/msc-report/raw/master/main.pdf}{Multisensor Architecture for an Intersection Management System}      
    \end{center}
\end{titlepage}