### 2018-08-17 main.tex

%\documentclass[draft]{book}
\documentclass{book}

\usepackage[utf8]{inputenc}
%\usepackage[spanish]{babel}
%\usepackage[spanish]{isodate}
\usepackage[final]{listings}
\usepackage{courier}
\usepackage{multirow}
\usepackage{enumitem}
% For links
\usepackage[bookmarks]{hyperref}
\hypersetup{colorlinks=true,allcolors=blue}

\usepackage[pdftex]{graphicx}
\usepackage[toc,page]{appendix}
\usepackage[nottoc]{tocbibind}

\usepackage{longtable}
\usepackage{pbox}
%\usepackage[sorting=none]{biblatex}
\usepackage{gensymb}
\usepackage{epigraph}

\usepackage{todonotes}

\usepackage{titling}
\usepackage{setspace}

\def\arraystretch{1.1}

%\renewcommand{\chaptername}{Capítulo}
%\renewcommand{\figurename}{Figure}

\lstset{	
	basicstyle=\footnotesize\ttfamily,
	language=bash,
	breaklines=true,
	fontadjust=true
}

\bibliographystyle{abbrv}
%\bibliographystyle{apalike}

\hyphenation{detection camera cameras Systems}

\newcommand\MyBox[2]{
  \fbox{\lower0.75cm
    \vbox to 1cm{\vfil
      \hbox to 1cm{\hfil\parbox{1.1cm}{#1\\#2}\hfil}
      \vfil}%
  }%
}

\begin{document}

\input{title1}
\input{title2}

\input{last_version}

%\title{Multisensor Architecture for an Intersection Management System\vspace{8cm}}
%\vfill
%\vfill
%\vfill
%\author{Gustavo Velasco-Hernandez\vspace{5cm}}
%\date{\today}
%\maketitle
%
%\title{Multisensor Architecture for an Intersection Management System}
%\vfill
%\vfill
%\vfill
%\author{Gustavo Velasco-Hernandez}
%\vfill
%\vfill
%\vfill
%\date{\today}
%\maketitle

\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}%
Intersections or junctions are critical points in transportation systems due to their dynamic behaviour, making them prone to safety or efficiency problems. For this reason the use of technology in monitoring intersections has increased over the last years. After a comprehensive review on this topic, a  multisensor architecture for an intersection management system is conceived, having as main features its scalability and modularity. This architecture is composed by four main elements: communication structure, data model, information structure and processing blocks. An example implementation of the architecture is made using the publisher/subscriber approach along with different tools and technologies that allow the whole system to fulfill the proposed criteria, to be scalable and modular. After testing the system using a dataset containing data from camera sensors and lasers sensors, it has been found that the use of heterogeneous data can lead to better results, at a low cost in complexity using the proposed framework. Comments about  obtained results, future developments and integration with current needs and new trends, are discussed.

%\chapter*{Acknowledgements}%
%\addcontentsline{toc}{chapter}{Acknowledgements}%

%I would like to thank to all people that in some way or another BLABLABLA

\tableofcontents
\setcounter{tocdepth}{3}

\listoffigures

\listoftables

\input{ch1-introduction}

\include{ch2-mdf_ims}

\include{ch3-arch_sys}

\include{ch4-results}

\include{ch5-conclusions}

%\begin{appendices}
%	\chapter {Vehicular Environment Simulators}
%\end{appendices}

\bibliography{bibliography}
%\bibliography{ref}

\end{document}