### 2019-01-22 last_version.tex
Removed TODOs and added revision history. Updated version

\begin{titlepage}
    \begin{center}
%        \vspace*{1cm}
%        \Huge        
%        Multisensor Architecture for an Intersection Management System
%   
%        \vspace{3.5cm}
%        \LARGE
%        Gustavo Adolfo Velasco-Hernandez
%        
%        \vfill
%        \large
%		        
%        
%        A thesis presented in partial fulfillment of the Requirement for the degree of Master of Science in Engineering with emphasis on Electronics.
%        
%        \vspace{2cm}
%        
%        Advisor: Prof. Eduardo Caicedo Bravo
%		
%		\vspace{2cm}
%        
%        %\includegraphics[width=0.4\textwidth]{university}
%        Perception and Intelligent Systems Research Group \\
%        School of Electrical and Electronics Engineering\\
%        Universidad del Valle\\
%        Colombia\\
%        \today
%\vfill
\todo{Update version message and link}
Version: 1.1 \\
\vspace{2cm}

\end{center}

Check for an updated version at github: 
\href{https://github.com/gustavovelascoh/msc-report/raw/master/main.pdf}{Multisensor Architecture for an Intersection Management System}

\subsection*{Revision history}
\subsubsection*{Version 1.1 - 2019-01-24:}

	\begin{itemize}
	\item Detailed processing levels in JDL model (1.1.1) and detailed categories in data fusion architectures (1.1.2)
	\item Fixed V2I description in figure 1.6 (1.2.1)
	\item Reworded description of processing stages in section 2.1.1
	\item Described three transportation simulation models (2.2.1)
	\end{itemize}

\subsubsection*{Version 1.0 - 2018-09-21:}
	\begin{itemize}
	\item Initial version
	\end{itemize}

\end{titlepage}